Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2021-01-17 minimax.tex
Some compression

%I find these settings useful in draft mode. Should be removed for final versions.
	%Which line breaks are chosen: accept worse lines, therefore reducing risk of overfull lines. Default = 200.
		\tolerance=2000
	%Accept overfull hbox up to...
		\hfuzz=2cm
	%Reduces verbosity about the bad line breaks.
		\hbadness 5000
	%Reduces verbosity about the underful vboxes.
		\vbadness=6000

@@

%This paper considers both sources of uncertainty at the same time, and proposes a method able to recommend a consensus choice without requiring full knowledge of either.
In this paper we focus on positional scoring rules, that are a particularly common method used to aggregate rankings, and we assume that both the agent preferences and the social choice rule are partially specified. 
We develop methods, based on the notion of minimax regret, for determining a robust ``winner'' under uncertainty of both the voting rule and the agent preferences.
%alternative using positional scoring rules and 
We provide incremental elicitation methods that 
at each step of the elicitation ask a question either to one of the agents or to the chair (the person or organization supervising the voting process) and discuss several heuristics that determine queries that quickly reduce regret. 
Answers to questions are encoded in terms of constraints; questions to the agents are comparions between pairs of alternatives while
questions to the chair ask  to select who should be the winner in a synthetic profile.
While previous works have considered either partial information about the agent preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.
Actually, very few works altogether have considered the problem of eliciting a voting rule by asking questions to the chair.

@@

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We consider a set $A$ of $m$ alternatives (products, restaurants, movies, public projects, job candidates, etc.) and a set $\set{1, …, n}$ of agents. Each agent $j$ comes from an infinite set $\N$ of potential agents and is associated to her “real” preference order ${\pref_j}  \in \linors$ which is a linear order (a connected, transitive, asymmetric relation) over the alternatives.
Following the social choice nomenclature, we call {\em profile} the association of a preference to each agent, considering a subset of agents from the set $\N$, and denote a profile by $(\pref_1,\ldots,\pref_n)$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

Let $V$ be the set of possible preference profiles (the union, for any integer $n$, of the $n$-fold Cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the empty set (sets are used for tied winners).
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$, also known as the {\em scoring vector}.
Let $\alpha^{x}_r$ be the number of times that alternative $x$ was ranked in the $r$-th position.
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
	\label{eq:srule}
	s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
	= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
The winners are the alternatives with the highest score.

In this work we assume fixed, but unknown to us, a profile $\profile^*$ representing the preferences $\pref_j^*$ of the agents, and a weight vector $\w^*$ representing the preferences of the chair. We will reason about partial preference information concerning those objects.
At a given time, our knowledge of agent $j$'s preference is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
%We use $\prefinc$ to denote incomparability, that is $a \prefinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each agent to a partial preference.

A completion of $\ppref_j$ is any linear order $\pref$ that extends $\ppref_j$ and we indicate with $C(\ppref_j) = \set{{\succ} \in \linors \suchthat {\ppref_j} \subseteq {\succ}}$ the set of possible completions of $\ppref_i$.
Then $C(\pprofile)=C(\ppref_1)\times … \times C(\ppref_n)$ is the set of complete profiles extending $p$. Note that $\profile^* \in C(\pprofile)$.

>>>>

%I find these settings useful in draft mode. Should be removed for final versions.
	%Which line breaks are chosen: accept worse lines, therefore reducing risk of overfull lines. Default = 200.
		\tolerance=2000
	%Accept overfull hbox up to...
		\hfuzz=2cm
	%Reduces verbosity about the bad line breaks. TODO reduce
		\hbadness 10000
	%Reduces verbosity about the underful vboxes.
		\vbadness=6000

@@

%This paper considers both sources of uncertainty at the same time, and proposes a method able to recommend a consensus choice without requiring full knowledge of either.
In this paper we focus on positional scoring rules, that are a particularly common method used to aggregate rankings. 
We develop methods, based on the notion of minimax regret, for determining a robust ``winner'' under uncertainty of both the voting rule and the agent preferences.
%alternative using positional scoring rules and 
We provide incremental elicitation methods that 
at each step of the elicitation ask a question either to one of the agents or to the chair (the person or organization supervising the voting process) and discuss several heuristics that determine queries that quickly reduce regret. 
Answers to questions are encoded in terms of constraints; questions to the agents are comparions between pairs of alternatives while
questions to the chair ask  to select who should be the winner in a synthetic profile.
While previous works have considered either partial information about the agent preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.
Actually, very few works altogether have considered the problem of eliciting a voting rule by asking questions to the chair.

@@

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We consider a set $A$ of $m$ alternatives (products, restaurants, public projects, job candidates, etc.) and a set $N = \set{1, …, n}$ of agents. Each agent $j$ comes from an infinite set $\N$ of potential agents and is associated to her “real” preference order ${\pref_j}  \in \linors$ which is a linear order (a connected, transitive, asymmetric relation) over the alternatives.
A {\em profile} $(\pref_1,\ldots,\pref_n)$ associates to each agent a preference, considering a subset of agents from the set $\N$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

Let $V$ be the set of possible preference profiles (the union, for any integer $n$, of the $n$-fold Cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates to each profile a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the empty set (sets are used for tied winners).
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$, also known as the {\em scoring vector}.
Let $\alpha^{x}_r$ be the number of times that alternative $x$ was ranked in the $r$-th position.
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
	\label{eq:srule}
	s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
	= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
The winners are the alternatives with the highest score.

We assume fixed, but unknown to us, a profile $\profile^*$ representing the preferences $\pref_j^*$ of the agents, and a weight vector $\w^*$ representing the preferences of the chair. 
At a given time, our knowledge of agent $j$'s preference is encoded by a partial order ${\ppref_j} \subseteq {\pref_j^*}$ over the alternatives, thus a transitive and asymmetric binary relation (we assume that preference information is truthful).
%We use $\prefinc$ to denote incomparability, that is $a \prefinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each agent to a partial preference.

Let $C(\ppref_j) = \set{{\succ} \in \linors \suchthat {\ppref_j} \subseteq {\succ}}$ denote the set of possible completions of $\ppref_i$ and $C(\pprofile) = \prod_{j \in N} C(\ppref_j)$ the set of complete profiles extending $\pprofile$. Note that $\profile^* \in C(\pprofile)$.
